---
title: "STA6933Final"
author: "Reverny Hsu, Noah Miller, Alexandra Perez"
date: "2025-04-21"
output: html_document
---

### Data Cleaning

```{r}
# Two separate data sets are available and are combined to have one larger dataset
d1 <- read.table("C:/Users/noahm/OneDrive/Documents/R/STA 6933/Final Project/CF_PCBA-686978.txt", sep="\t", header=T)
d2 <- read.table("C:/Users/noahm/OneDrive/Documents/R/STA 6933/Final Project/PCBA-686978.txt", sep="\t", header=T)
```

```{r}
# Combine two tables
data <- rbind(d1, d2)
```

```{r}
summary(data)
```

```{r}
# remove missing values
data <- na.omit(data)

# remove duplicated
library(dplyr)

data <- data %>%
  distinct()
```

```{r}
str(data)
table(data$Label)
```

The label variable is imbalance.

### EDA

```{r}
library(ggplot2)
library(corrplot)
library(reshape2)
library(GGally)

#Bar plot to show distribution of 'Label'
ggplot(data, aes(x = as.factor(Label))) +
  geom_bar(fill = "orange") +
  labs(x = "Label", y = "Count", title = "Class Distribution")


#Get numeric columns
numeric_cols <- names(data)[sapply(data, is.numeric)]
numeric_cols <- setdiff(numeric_cols, "Label")

#Histograms
par(mfrow = c(2, 3))
for (col in numeric_cols) {
  hist(data[[col]], main = col, xlab = col)
}

# Boxplots grouped by 'Label'
par(mfrow = c(2, 3))
for (col in numeric_cols) {
  boxplot(data[[col]], main = col)}
  
#T-test   
for (col in numeric_cols) {
  cat(col, "\n")
  print(t.test(data[[col]] ~ data$Label))
}

#Density Plots
for (col in numeric_cols) {
  p <- ggplot(data, aes_string(x = col, fill = "as.factor(Label)")) + 
    geom_density(alpha = 0.5) +
    ggtitle(paste("Density Plot of", col)) +
    theme_minimal()
  
  print(p)
}

#Correlation Graphs

plot_cols <- c(numeric_cols, "Label")

chunk_size <- 4
chunks <- split(plot_cols[plot_cols != "Label"], ceiling(seq_along(plot_cols[plot_cols != "Label"]) / chunk_size))

for (i in seq_along(chunks)) {
  chunk <- chunks[[i]]
  cols_to_plot <- c(chunk, "Label")
  
  p <- ggpairs(data[, cols_to_plot], aes(color = as.factor(Label)))
  print(p)
}

# Correlation plots
par(mfrow = c(1, 1))
corrplot(cor(data[, plot_cols]), method = "number", type = "upper")
```

### Scaling/Normalization & Handle imbalanced Data

```{r}
library(readr)

y <- data$Label

# Extract only numeric predictors (excluding Label)
x_scaled <- data[, sapply(data, is.numeric)]
x_scaled <- x_scaled[, setdiff(names(x_scaled), "Label")]

# Normalize
x_scaled <- as.data.frame(scale(x_scaled))
```

```{r}
library(smotefamily)
# Apply adas with desired parameters
adas <-ADAS(x_scaled, y, K = 5)
```

We ended up using ADAS to control imbalance data instead of using SMOTE, since:

- the mild class imbalance

- SMOTE doesn’t generate samples (especially with dup_size < 1)

ADAS is Better with mild imbalance, it works even without large oversampling.

```{r}
balanced_data <- adas$data

table(balanced_data$class)
```

The data is balance now ("0": 531; "1":528)


```{r}
# This splits the data into training and testing data sets.
set.seed(123)
train = sample(nrow(balanced_data), 0.7*nrow(balanced_data))
pfas.train = balanced_data[train,]
pfas.test = balanced_data[-train,]
```


# Logistic regression
```{r}
pfas.train$class <- as.numeric(pfas.train$class) 
pfas.test$class <- as.numeric(pfas.test$class) 

# Logistic Regression Model
lg_model <- glm(class ~ ., data = pfas.train, family = binomial)
summary(lg_model)
```

```{r}
# Make predictions on the test data
pred <- predict(lg_model, pfas.test, type = "response")

# Convert predictions to binary outcomes (0 or 1) based on a threshold of 0.5
pred_binary<- ifelse(pred > 0.5, 1, 0)
```

```{r}
# performance metrics
library(pROC)
library(caret)

confusion_matrix <- table(pfas.test$class, pred_binary)

# F1-score calculation
precision <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])
recall <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])
f1_score <- 2 * (precision * recall) / (precision + recall)
print(paste("F1-score: ", f1_score))

# MCC (Matthews Correlation Coefficient) calculation
TP <- confusion_matrix[2, 2]
TN <- confusion_matrix[1, 1]
FP <- confusion_matrix[1, 2]
FN <- confusion_matrix[2, 1]
mcc <- (TP * TN - FP * FN) / sqrt((TP + FP) * (TP + FN) * (TN + FP) * (TN + FN))
print(paste("MCC: ", mcc))
```

```{r}
# AUC-ROC calculation
roc_curve <- roc(pfas.test$class, pred) 

# Plot ROC curve
plot(roc_curve, col = "blue", main = "ROC Curve - Logistic Regression")

# Calculate AUC-ROC
auc_value <- auc(roc_curve)
print(paste("AUC-ROC: ", auc_value))
```

```{r}
# Logistic Regression Model w/ only significant predictors
lg_model2 <- glm(class ~ Estimateddensity + Overallaveragevalue + Positiveaveragevalue + Molecularpolarityindex, data = pfas.train, family = binomial)
summary(lg_model2)
```

```{r}
# Make predictions on the test data
pred <- predict(lg_model2, pfas.test, type = "response")

# Convert predictions to binary outcomes (0 or 1) based on a threshold of 0.5
pred_binary<- ifelse(pred > 0.5, 1, 0)
```

```{r}
# performance metrics
library(pROC)
library(caret)

confusion_matrix <- table(pfas.test$class, pred_binary)

# F1-score calculation
precision <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])
recall <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])
f1_score <- 2 * (precision * recall) / (precision + recall)
print(paste("F1-score: ", f1_score))

# MCC (Matthews Correlation Coefficient) calculation
TP <- confusion_matrix[2, 2]
TN <- confusion_matrix[1, 1]
FP <- confusion_matrix[1, 2]
FN <- confusion_matrix[2, 1]
mcc <- (TP * TN - FP * FN) / sqrt((TP + FP) * (TP + FN) * (TN + FP) * (TN + FN))
print(paste("MCC: ", mcc))
```

```{r}
# AUC-ROC calculation
roc_curve <- roc(pfas.test$class, pred) 

# Plot ROC curve
plot(roc_curve, col = "blue", main = "ROC Curve - Logistic Regression")

# Calculate AUC-ROC
auc_value <- auc(roc_curve)
print(paste("AUC-ROC: ", auc_value))
```




```{r}
# Logistic Regression Model w/ polynomial
lg_model3 <- glm(class ~ Volume + Estimateddensity + Minimal+ poly(Positivevariance,2) + Overallaveragevalue+ Positiveaveragevalue+ Balanceofcharges + poly(Maximal,2) + poly(Overallvariance,2) + Product + Molecularpolarityindex, data = pfas.train, family = binomial)
summary(lg_model3)
```

```{r}
# Make predictions on the test data
lg_preds <- predict(lg_model3, pfas.test, type = "response")

# Convert predictions to binary outcomes (0 or 1) based on a threshold of 0.5
pred_binary<- ifelse(pred > 0.5, 1, 0)
```
```{r}
# AUC-ROC calculation
roc_curve_lg  <- roc(pfas.test$class, lg_preds) 

# Plot ROC curve
plot(roc_curve_lg, col = "blue", main = "ROC Curve - Logistic Regression")

# Calculate AUC-ROC
auc_value_lg  <- auc(roc_curve_lg)
print(paste("AUC-ROC: ", auc_value_lg))
```

```{r}
confusion_matrix <- table(pfas.test$class, pred_binary)

# F1-score calculation
precision <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])
recall <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])
f1_score <- 2 * (precision * recall) / (precision + recall)
print(paste("F1-score: ", f1_score))

# MCC (Matthews Correlation Coefficient) calculation
TP <- confusion_matrix[2, 2]
TN <- confusion_matrix[1, 1]
FP <- confusion_matrix[1, 2]
FN <- confusion_matrix[2, 1]
mcc <- (TP * TN - FP * FN) / sqrt((TP + FP) * (TP + FN) * (TN + FP) * (TN + FN))
print(paste("MCC: ", mcc))
```
The AUC-ROC value for this logistic regression model is the highest of the models, which makes it the best one to use.


# Forward Stepwise Selection
```{r}
library(leaps)
pfas.forward <- regsubsets(class ~ ., data = pfas.train, nvmax = 18, method = "forward")
pfas_sum <- summary(pfas.forward)
```
```{r}
# These plots find the number of predictors that are best to use.
par(mfrow=c(2,2))
plot(pfas_sum$rsq ,xlab="Number of Variables ",ylab="RSq", type='l')
index = which.min(pfas_sum$rsq) # 1 predictors
points(index,pfas_sum$rsq[index],col="red",cex=2,pch=20)

plot(pfas_sum$bic ,xlab="Number of Variables ",ylab="BIC",type='l')
index = which.min(pfas_sum$bic) # 4 predictors
points(index,pfas_sum$bic[index],col="red",cex=2,pch=20)

plot(pfas_sum$adjr2 ,xlab="Number of Variables ", ylab="Adjusted RSq",type="l")
index = which.max(pfas_sum$adjr2) # 9 predictors
points(index,pfas_sum$adjr2[index], col="red",cex=2,pch=20)
```
```{r}
coef(pfas.forward, 4)
```

# XGBoost
```{r}
library(xgboost)
dtrain <- xgb.DMatrix(data = as.matrix(sapply(pfas.train[, -ncol(pfas.train)], as.numeric)), label = as.numeric(pfas.train$class))
xg.pfas <- xgboost(data = dtrain, nrounds = 100, objective = "binary:logistic")
xg.pfas.imp <- xgb.importance(model = xg.pfas)
xgb.plot.importance(xg.pfas.imp, rel_to_first = TRUE, xlab = "Relative importance")
```

```{r}
# XGBoost w/ hyperparameter n_estimators=200, learning_rate=0.1, max_depth=20
set.seed(42)
library(caret)      # For F1 and MCC calculation
library(pROC)       # For ROC AUC calculation

# Prepare the data
dtrain <- xgb.DMatrix(data = as.matrix(sapply(pfas.train[, -ncol(pfas.train)], as.numeric)), 
                      label = as.numeric(pfas.train$class))
dtest <- xgb.DMatrix(data = as.matrix(sapply(pfas.test[, -ncol(pfas.test)], as.numeric)), 
                     label = as.numeric(pfas.test$class))

# Define parameters for XGBoost model
params <- list(
  objective = "binary:logistic",   # binary classification problem
  eta = 0.01,                       # learning rate
  max_depth = 25,                   # maximum depth of trees
  colsample_bytree = 0.7,
  gamma=0,
  min_child_weight = 1,
  nthread = 2,                      # number of threads to use
  eval_metric = "error"             # evaluation metric
)

# Train the model
xg.pfas2 <- xgboost(params = params, 
                   data = dtrain, 
                   nrounds = 300,  # number of boosting rounds
                   verbose = 1)

# Make predictions on the test set (probabilities)
xg_preds2 <- predict(xg.pfas2, dtest)

# Convert probabilities to binary outcomes using a threshold of 0.5
pred_binary <- ifelse(xg_preds2 > 0.5, 1, 0)

# Calculate confusion matrix
conf_matrix <- table(pfas.test$class, pred_binary)

# F1-score Calculation
precision <- conf_matrix[2, 2] / sum(conf_matrix[2, ])
recall <- conf_matrix[2, 2] / sum(conf_matrix[, 2])
f1_score <- 2 * (precision * recall) / (precision + recall)
print(paste("F1-score: ", f1_score))

# MCC (Matthews Correlation Coefficient) Calculation
TP <- conf_matrix[2, 2]
TN <- conf_matrix[1, 1]
FP <- conf_matrix[1, 2]
FN <- conf_matrix[2, 1]
mcc <- (TP * TN - FP * FN) / sqrt((TP + FP) * (TP + FN) * (TN + FP) * (TN + FN))
print(paste("MCC: ", mcc))

# ROC AUC Calculation
roc_curve_boost2 <- roc(pfas.test$class, xg_preds2)  # Use predicted probabilities for ROC
auc_value_boost2 <- auc(roc_curve_boost2)
print(paste("ROC AUC: ", auc_value_boost2))

# Plot ROC curve
#plot(roc_curve, col = "blue", main = "ROC Curve - XGBoost")
#text(x = 0.6, y = 0.4, labels = paste("AUC = ", round(auc_value, 3)), col = "red", cex = 1.2)
```

# Random Forest
```{r}
library(randomForest)
pfas.forest <- randomForest(as.factor(class) ~ ., data = pfas.train, importance = TRUE, ntree = 5000)
varImpPlot(pfas.forest)
```

# Random Forest w/ tuning hyperparameters
```{r}
library(MLmetrics)
# Train the random forest model with the specified parameters
library(randomForest)
library(pROC)
set.seed(42)
# Train the random forest model
pfas.forest2 <- randomForest(
  as.factor(pfas.train$class) ~ ., 
  data = pfas.train, 
  importance = TRUE, 
  ntree = 200,             # Number of trees in the forest
  max_depth = 20,          # Limiting max depth (control tree depth)
  min_samples_leaf = 5,
  min_samples_split = 1,
  nodesize = 5,            # Minimum size of nodes
  mtry = sqrt(ncol(pfas.train) - 1)  # Default mtry, can be tuned
)

# Make predictions on the test set
pred_binary <- predict(pfas.forest2, pfas.test)

# Confusion matrix for calculating F1 and MCC
conf_matrix <- table(pfas.test$class, pred_binary)

# F1-score Calculation
precision <- conf_matrix[2, 2] / sum(conf_matrix[2, ])
recall <- conf_matrix[2, 2] / sum(conf_matrix[, 2])
f1_score <- 2 * (precision * recall) / (precision + recall)
print(paste("F1-score: ", f1_score))

# MCC (Matthews Correlation Coefficient) Calculation
TP <- conf_matrix[2, 2]
TN <- conf_matrix[1, 1]
FP <- conf_matrix[1, 2]
FN <- conf_matrix[2, 1]
mcc <- (TP * TN - FP * FN) / sqrt((TP + FP) * (TP + FN) * (TN + FP) * (TN + FN))
print(paste("MCC: ", mcc))

# ROC AUC Calculation
for_preds2 <- predict(pfas.forest2, pfas.test, type = "prob")  # Get probabilities for ROC
roc_curve_for2 <- roc(pfas.test$class, as.numeric(for_preds2[, 2]))  # Use probabilities for the positive class
auc_value_for2 <- auc(roc_curve_for2)
print(paste("ROC AUC: ", auc_value_for2))
```

# SVM
```{r}
library(e1071)

# Separate features and labels for training and testing
X_train <- pfas.train[, -which(names(pfas.train) == "class")]  # All columns except 'class'
y_train <- pfas.train$class  # Assuming 'class' is the target variable
X_test <- pfas.test[, -which(names(pfas.test) == "class")]
y_test <- pfas.test$class

# Train the SVM model with RBF kernel
svm_model <- svm(
  as.factor(y_train) ~ ., 
  data = cbind(X_train, y_train), 
  kernel = "radial",  # RBF kernel
  cost = 100,         # C = 100 (Regularization parameter)
  probability = TRUE, # To get probability estimates
  random_state = 42,
  sigma = 0.01
)

# Make predictions on the test set
pred_binary <- predict(svm_model, pfas.test[, -which(names(pfas.test) == "class")])
pred_svm <- attr(predict(svm_model, pfas.test[, -which(names(pfas.test) == "class")], probability = TRUE), "probabilities")[, 2] # Probability of positive class

# Confusion matrix for calculating F1 and MCC
conf_matrix <- confusionMatrix(factor(pred_binary), factor(y_test))
print(conf_matrix)

# MCC (Matthews Correlation Coefficient) Calculation
TP <- conf_matrix$table[2, 2]
TN <- conf_matrix$table[1, 1]
FP <- conf_matrix$table[1, 2]
FN <- conf_matrix$table[2, 1]
mcc <- (TP * TN - FP * FN) / sqrt((TP + FP) * (TP + FN) * (TN + FP) * (TN + FN))
print(paste("MCC: ", mcc))

# F1-score Calculation
precision <- conf_matrix$byClass["Pos Pred Value"]
recall <- conf_matrix$byClass["Sensitivity"]
f1_score <- 2 * (precision * recall) / (precision + recall)
print(paste("F1-score: ", f1_score))

# ROC AUC Calculation
roc_curve_svm <- roc(pfas.test$class, pred_svm)  # Use predicted probabilities for ROC curve
auc_value_svm <- auc(roc_curve_svm)
print(paste("ROC AUC: ", auc_value_svm))
```

# Light GBM
```{r}
library(lightgbm)

# Prepare the data
X_train <- pfas.train[, -which(names(pfas.train) == "class")]  # Features
y_train <- pfas.train$class 

X_test <- pfas.test[, -which(names(pfas.test) == "class")]
y_test <- pfas.test$class

# Convert data to LightGBM dataset format
train_data <- lgb.Dataset(data = as.matrix(X_train), label = y_train)
test_data <- lgb.Dataset(data = as.matrix(X_test), label = y_test)

# Set parameters for LightGBM
params <- list(
  objective = "binary",          # For binary classification
  metric = "auc",                # Use AUC for evaluation
  boosting_type = "gbdt",        # Gradient Boosting Decision Tree
  num_leaves = 31,               # Number of leaves in one tree
  learning_rate = 0.01,          # Learning rate
  feature_fraction = 0.8,         # Randomly select a subset of features
  bagging_fraction = 0.8,
  max_depth = 15,
  min_data_in_leaf = 20
)


valid = list(test = test_data)
#model training
lt_gbm_model = lightgbm(params = params, train_data, valid, nrounds = 1000)

# Make predictions on the test set
lt_gbm_pred <- predict(lt_gbm_model, as.matrix(pfas.test[, -which(names(pfas.test) == "class")]))

# Convert probabilities to binary predictions (threshold = 0.5)
y_pred_binary <- ifelse(lt_gbm_pred > 0.5, 1, 0)  # For binary classification

# Confusion matrix
conf_matrix <- confusionMatrix(factor(y_pred_binary), factor(y_test))
print(conf_matrix)

# Calculate F1-score
precision <- conf_matrix$byClass["Pos Pred Value"]
recall <- conf_matrix$byClass["Sensitivity"]
f1_score <- 2 * (precision * recall) / (precision + recall)
print(paste("F1-score: ", f1_score))

# Calculate ROC AUC
roc_curve_ltgbm <- roc(pfas.test$class, lt_gbm_pred)  # Using predicted probabilities for ROC curve
auc_value_ltgbm <- auc(roc_curve_ltgbm)
print(paste("ROC AUC: ", auc_value_ltgbm))
```

```{r}
# Extract confusion matrix values
TP <- conf_matrix$table[2, 2]  # True Positives
TN <- conf_matrix$table[1, 1]  # True Negatives
FP <- conf_matrix$table[1, 2]  # False Positives
FN <- conf_matrix$table[2, 1]  # False Negatives

# Calculate Matthews Correlation Coefficient (MCC)
mcc <- (TP * TN - FP * FN) / sqrt((TP + FP) * (TP + FN) * (TN + FP) * (TN + FN))

# Print MCC value
print(paste("MCC: ", mcc))
```


# BART Model
```{r}
pfas.train[] <- lapply(pfas.train, function(x) if (is.character(x)) as.factor(x) else x)
pfas.test[]  <- lapply(pfas.test,  function(x) if (is.character(x)) as.factor(x) else x)

library(BART)
set.seed(34)
pfas.bart <- pbart(as.data.frame(pfas.train[,-ncol(pfas.train)]), pfas.train$class, as.data.frame(pfas.test[,-ncol(pfas.test)]))
bartImp <- colMeans(pfas.bart$varcount)
head(sort(bartImp, decreasing = T), 10)
```

# GAM Model
```{r}
library(gam)
library(mgcv)
library(splines)
par(mfrow=c(2,2))
gam.pfas <- gam(class ~ s(Volume) + s(Estimateddensity) + s(Negativevariance) + s(Overallaveragevalue), family = binomial, data = pfas.train)
plot(gam.pfas, se = TRUE, col = "blue")
```

# Stacked models
```{r}
# XGBoost
xg_preds2 <- predict(xg.pfas2, dtest)

# Random Forest
rf_preds <- predict(pfas.forest2, pfas.test, type = "response")

# SVM
svm_preds <- attr(predict(svm_model, pfas.test[, -which(names(pfas.test) == "class")], probability = TRUE), "probabilities")[, 2]

# Stack the predictions 
stacked_data <- data.frame(xgboost = xg_preds2, random_forest = rf_preds, svm = svm_preds)

# Train the meta-model using the stacked predictions
meta_model <- glm(y_test ~ ., data = stacked_data, family = binomial)

# Make predictions with the stacked model
stacked_preds <- predict(meta_model, newdata = stacked_data, type = "response")

roc_curve <- roc(y_test, stacked_preds)
auc(roc_curve)
```

```{r}
stacked_binary_preds <- ifelse(stacked_preds > 0.5, 1, 0)
conf_matrix <- table(Predicted = stacked_binary_preds, Actual = y_test)

# F1 score calculation
precision <- conf_matrix[2, 2] / sum(conf_matrix[2, ])
recall <- conf_matrix[2, 2] / sum(conf_matrix[, 2])
f1_score <- 2 * (precision * recall) / (precision + recall)
cat("F1 Score:", f1_score, "\n")

# MCC calculation
mcc <- (conf_matrix[1, 1] * conf_matrix[2, 2] - conf_matrix[1, 2] * conf_matrix[2, 1]) / 
      sqrt((conf_matrix[1, 1] + conf_matrix[1, 2]) * (conf_matrix[1, 1] + conf_matrix[2, 1]) * 
           (conf_matrix[2, 2] + conf_matrix[1, 2]) * (conf_matrix[2, 2] + conf_matrix[2, 1]))
cat("MCC:", mcc, "\n")
```


# Hierarchical clustering with feature importance
```{r}
library(ggplot2)
library(cluster)

importance_df <- data.frame(Feature = rownames(importance(pfas.forest)),
                            Importance = importance(pfas.forest)[, "MeanDecreaseGini"])

# Sort by importance
importance_df <- importance_df %>% arrange(desc(Importance))
print(importance_df)
```
```{r}
top_features <- importance_df$Feature[1:5]
X_top <- balanced_data %>% select(all_of(top_features))

distance_matrix <- dist(X_top)

# Hierarchical clustering using the distance matrix
hclust_model <- hclust(distance_matrix, method = "ward.D2")
```

```{r}
# Cut the tree into clusters
k <- 5
clusters <- cutree(hclust_model, k)

# Attach clusters back to the original data
balanced_data$cluster <- as.factor(clusters)
```

```{r}
# Visualize Clusters vs Important Descriptors

# Positive Surface Area vs Clusters
ggplot(balanced_data, aes(x = cluster, y = Positivesurfacearea, fill = cluster)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Positive Surface Area across Clusters", y = "Positive Surface Area", x = "Cluster")

# Summarize cluster means
cluster_summary <- balanced_data %>%
  group_by(cluster) %>%
  summarise(across(all_of(top_features), mean, na.rm = TRUE))

print(cluster_summary)
```

```{r}
library(dplyr)

cluster_summary <- balanced_data %>%
  group_by(cluster) %>%
  summarise(
    total = n(),
    active = sum(class == 1),
    active_rate = mean(class == 1)
  ) %>%
  arrange(desc(active_rate))

print(cluster_summary)
```

```{r}
ggplot(cluster_summary, aes(x = cluster, y = active_rate, fill = cluster)) +
  geom_bar(stat = "identity") +
  labs(title = "Bioactivity Rate per Cluster",
       x = "Cluster",
       y = "Proportion of Active (class = 1)") +
  theme_minimal()
```

```{r}
library(dplyr)

descriptor_summary <- balanced_data %>%
  group_by(cluster) %>%
  summarise(
    ActiveRate = mean(class == 1),
    Estimateddensity = mean(Estimateddensity, na.rm = TRUE),
    Positivesurfacearea = mean(Positivesurfacearea, na.rm = TRUE),
    Volume = mean(Volume, na.rm = TRUE)
  )

print(descriptor_summary)
```

```{r}
ggplot(balanced_data, aes(x = cluster, y = Estimateddensity, fill = cluster)) +
  geom_boxplot() +
  labs(title = "Estimateddensity by Cluster", y = "Estimateddensity") +
  theme_minimal()
```

```{r}
ggplot(balanced_data, aes(x = cluster, y = Positivesurfacearea, fill = cluster)) +
  geom_boxplot() +
  labs(title = "Positivesurfacearea by Cluster", y = "Positivesurfacearea") +
  theme_minimal()
```

```{r}
ggplot(balanced_data, aes(x = cluster, y = Volume, fill = cluster)) +
  geom_boxplot() +
  labs(title = "Volume by Cluster", y = "Volume") +
  theme_minimal()
```

# ROC-AUC
```{r}
library(pROC)
test_features <- as.matrix(sapply(pfas.test[, -ncol(pfas.test)], as.numeric))

#roc for logistic regression w/ poly
lg_preds <- predict(lg_model3, pfas.test, type = "response")
roc_curve_lg <- roc(pfas.test$class, lg_preds) 
auc_value_lg <- auc(roc_curve_lg)

#roc for xgboost
xg_preds2 <- predict(xg.pfas2, dtest)
roc_curve_boost2 <- roc(pfas.test$class, xg_preds2)
auc_value_boost2 <- auc(roc_curve_boost2)

#roc for random forest
for_preds2 <- predict(pfas.forest2, pfas.test, type = "prob") 
roc_curve_for2 <- roc(pfas.test$class, as.numeric(for_preds2[, 2]))  
auc_value_for2 <- auc(roc_curve_for2)

#roc for SVM
pred_svm <- attr(predict(svm_model, pfas.test[, -which(names(pfas.test) == "class")], probability = TRUE), "probabilities")[, 2]
roc_curve_svm <- roc(pfas.test$class, pred_svm)  
auc_value_svm <- auc(roc_curve_svm)

#roc for lightGBM
lt_gbm_pred <- predict(lt_gbm_model, as.matrix(pfas.test[, -which(names(pfas.test) == "class")]))
roc_curve_ltgbm <- roc(pfas.test$class, lt_gbm_pred)
auc_value_ltgbm <- auc(roc_curve_ltgbm)

#roc for bart model
roc_curve_bart <- roc(pfas.test$class, pfas.bart$prob.test.mean)
auc_value_bart <- auc(roc_curve_bart)

#roc for GAM Model
gam_preds <- predict(gam.pfas, newdata = pfas.test, type = "response")
roc_curve_gam <- roc(pfas.test$class, gam_preds)
auc_value_gam <- auc(roc_curve_gam)

#roc for stack model
stacked_preds <- predict(meta_model, newdata = stacked_data, type = "response")
roc_curve_stacked <- roc(y_test, stacked_preds)
auc_value_stacked <- auc(roc_curve_stacked)

# The following lines created the plots for the ROC curves
plot(roc_curve_boost2, col = "blue", main = "ROC Curves for Different Models", 
     xlab = "False Positive Rate", ylab = "True Positive Rate", lwd = 2)
plot(roc_curve_for2, col= "green", add=TRUE, lwd=2)
plot(roc_curve_lg, col = "orange", add = TRUE, lwd = 2)
plot(roc_curve_svm, col = "pink", add = TRUE, lwd = 2)
plot(roc_curve_bart, col = "purple", add = TRUE, lwd = 2)
plot(roc_curve_gam, col = "red", add = TRUE, lwd = 2)
plot(roc_curve_ltgbm, col = "yellow", add = TRUE, lwd = 2)
plot(roc_curve_stacked, col = "black", add = TRUE, lwd = 2)

legend("bottomright", legend = c(paste("XGBoost:", round(auc_value_boost2, 4)),
                                 paste("Random Forest:", round(auc_value_for2, 4)),
                                 paste("Logistic Regression:", round(auc_value_lg, 4)),
                                 paste("SVM:", round(auc_value_svm, 4)),
                                 paste("BART:", round(auc_value_bart, 4)),
                                 paste("GAM Model:", round(auc_value_gam, 4)),
                                 paste("Light GBM Model:", round(auc_value_ltgbm, 4)),
                                 paste("Stacked Model:", round(auc_value_stacked, 4))),
       col = c("blue", "green", "orange", "pink", "purple", "red","yellow","black"), lwd = 2)
```

